---
title: "Assignment2"
author: "Lina Barbosa"
date: "5/10/2017"
output: 
  pdf_document: 
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# 1. Loading Data

Read dataset as tibble

```{r}
gaz_raw <- read_delim("./CA_Features_20170401.zip", delim ="|", col_types = cols()) %>% 
  as_tibble()
```

# 2. Questions 1-4

I created gaz tibble by selecting the listed variables. I changed the names of the variables to lower case, and formatted properly the Date_created and Date_edited varibles as.Date. Also, I changed the Unknown values to NAs and got rid of of the 0 values in the variables longitude and latitude.

```{r}
gaz <- gaz_raw %>% 
  select(FEATURE_ID = 1,
         starts_with("FEATURE"),
         STATE_ALPHA,
         COUNTY_NAME,
         ends_with("DEC"),
         ELEV_IN_M,
         MAP_NAME,
         starts_with("DATE")) %>% 
  magrittr::set_colnames(value = tolower(colnames(.))) %>% 
  mutate(date_created = as.Date(date_created, format = "%m/%d/%y"),
         date_edited = as.Date(date_edited, format = "%m/%d/%y"),
         map_name = ifelse(map_name == "Unknown", NA, map_name)) %>% 
  filter(prim_lat_dec != 0, prim_long_dec != 0, state_alpha == "CA")

# This is not part of the assignment (just me having fun) 
# check_col_vals <- function(data){
#   sapply(data, function(x){sort(unique(x))})
# }

```

# Question 5

Exporting gaz tibble as csv 

```{r}
write_delim(gaz, path = "./gaz.csv", delim = "|")
```

# Analyze 

## 1. What is the most-frequently-occuring feature name?

```{r}
counts_name <- gaz %>% 
  group_by(feature_name) %>% 
  summarise(count = n()) %>% 
  arrange(count)

# I'm using this -- counts_name$feature_name[counts_name$count == max(counts_name$count)] 
# -as an inline code below

```

The most-frequently-occuring feature name `r counts_name$feature_name[counts_name$count == max(counts_name$count)]`.

## What is the least-frequently-occuring feature class?

```{r}
counts_class <- gaz %>% 
  group_by(feature_class) %>% 
  summarise(count = n()) %>% 
  arrange(count)

```

The least-frequently-occuring feature class is `r counts_class$feature_class[counts_class$count == min(counts_class$count)]`

## 3. What is the approximate center point of each county?


```{r}
center_point <- gaz %>% 
  group_by(county_name) %>% 
  summarise(latitude = mean(prim_lat_dec, na.rm = TRUE),
            longitude = mean(prim_long_dec, na.rm = TRUE),
            longmax = max(prim_long_dec, na.rm = TRUE),
            longmin = min(prim_long_dec, na.rm = TRUE),
            latmax = max(prim_lat_dec, na.rm = TRUE),
            latmin = min(prim_lat_dec, na.rm = TRUE),
            lat = (latmax + latmin)/2,
            long = (longmax + longmin)/2) %>% 
  select(-c(longmax, latmax, longmin, latmin))
            
```

```{r, fig.cap="Black points represent the center point calculated using the mean of the latitude and longitud points for each county. Blue points also represent the center point of the bounding box"}

ggplot(gaz, aes(x = prim_long_dec, y = prim_lat_dec, color = county_name)) +
  geom_point(pch = ".") +
  geom_point(data = center_point, aes(x = longitude, y = latitude), color = "black") +
  geom_point(data = center_point, aes(x = long, y = lat), color = "darkblue") +
  coord_quickmap() +
  theme_bw() +
  theme(legend.position = "none")
```

```{r}
# FC <- gaz %>% 
#   select(feature_class) %>% 
#   unique()

# write_csv(FC, "FC.csv")
```

# 4. What are the fractions of the total number of features in each county that are natural? man-made?

```{r}
FC2 <- read_csv("FC.csv", col_types = cols())

class_made <- gaz %>% 
  left_join(FC2) %>% 
  group_by(county_name, made_by) %>% 
  summarise(count = n()) %>% 
  spread(made_by, count) %>% 
  mutate(propman = man/(man + nature),
         propnature = 1-propman) 

write_csv(class_made, "Question4.csv")
```

The file "Question4.csv" contains the proportion of the total number of features in each county that are natural and man-made